---
title: ' '
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{float}
- \usepackage{sectsty}
---


# Introduction

## Overview to chosen Approach 

Insurers provide cover under adverse events for their clients. Assessing risk premiums within respective tariff classes is of key issue for actuaries to ensure that insurance claims can be served. \citep{antonio_iabe_nodate,robben_assessing_2022} This work provides an analysis on the comparison of different modelling approaches to risk parameters in the case of automobile insurance, building a structure of proposed tariff groups with respectively calculated risk loadings on top of it. The methodic chapter is split into two steps due to this reason. To a certain degree the general structure of this scientific work is followed. \citep{antonio_iabe_nodate,robben_assessing_2022} In the first part we estimate the pure premium by modelling claim frequency and severity seperately to then evaluate the premium. Within  modeling we considered GLM and XGboost models and evaluated their performance to solve the task at hand. Finally in a second stage we cluster clients into groups of similar risk and  provide a method to calculate risk loadings on the group level of our tariff structure. 


## Exploration of the Motor Third Party Liability Dataset

The data at hand is the Motor Third Party Liability Dataset. It is two dimensional structured with columns specifying features for individual clients indexed by rows. There are 20 different features and 163,647 individual rows. The following table from the assignment task identifies the short Feature names. Our Target variables for modeling are nbrtotc, the total number of claims during exposure and the severity which is calculated as $sev = chargtot / nbrtotc$. The variables which are most important for the explanatory modelling are almost all categorical or ordinal features, which makes this part somewhat easier. Only the Age of Policy Holder and Coordinates are continuous features useful to explain $nbrtotc$ and $sev$. 


```{r, fig.align='center',fig.cap="Feature Explanation",  echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)

Feature <- c("ageph", "CODPOSS", "duree", "lnexpo", "nbrtotc", "chargtot", "agecar", 
           "sexp", "fuelc", "split", "usec", "fleetc", "sportc", "coverp", "powerc")

Explanation <- c("age of the policyholder", 
  "postal code in Belgium", 
  "exposure, fraction of the year the insured is covered",
  "log of exposure", 
  "total number of claims during period of exposure",
  "total claim amount",
  "ageofthecar: 0−1,2−5,6−10,>10", 
  "sex of the policyholder: male or female",
  "type of fuel: petrol or gasoil",
  "split of the premium: monthly, once, twice, three times per year",
  "use of the car: private or professional",
  "car belonging to a fleet: yes or no", 
  "sport car: yes or no",
  "coverage: MTPL, MTPL+, MTPL+++",
  "power of the car: < 66, 66-110, >110")

frame <- data.frame(Feature, Explanation)

knitr::kable(frame, "simple")

```


Furthermore the postal code gives us information about where in Belgium the policy holder resides. The following plot Figure 1: Spatial Exposure additionally highlights how many policyholder in out set reside in which regions. Metropolitan areas and large cities are in visible contrast to less populated zones. Especially the southern part of Belgium closer to the french border, hosts very few observations. This map is of course slightly similar to the population distribution in Belgium.  
  
  
```{r fig.align='center',fig.width=5, fig.height=4, echo=FALSE, fig.cap="Spatial Exposure", warning=FALSE, message=FALSE}
source("01_dataprep.R")
be_expo_reg

```


\newpage
Taking a closer look at the frequency of certain features reveals interesting insights about our dataset. Starting with the first plot in the top left of Figure 2: Summary Plots we take a look at $duree$, it shows that most policy holder were insured during the whole year. The histograms for $nbrtotan$ and $nbrtotc$ reveal then again what we know from academic theory: the absolute majority of policy holders have zero claims.The age of the car was binned into groups prior and shows that intervals with higher car ages become larger and then drop off for the highest group which includes car which are older than 10 years. Most individuals in our set pay once a year, use the car in a private context, which is not in a fleet and not a sportscar. Given that sport cars are not really represented in the data it makes sense that almos all cars have less than 110 hp. 


```{r fig.align='center',fig.width=8, fig.height=10, echo=FALSE, fig.cap="Summary Plots", warning=FALSE, message=FALSE}

grid.arrange(grobs = g_list, ncol = 3)
```

The distribution of features is very insightful into the composition of our data. Next up we would like to take a closer look at the interaction between certain variables an our target variables $nbrtotc$ and $sev$, which will be very important for the interpretation of our modeling section. 






 
